Population Differentiation Analysis

### 2017-07-06 doc/RMD/pop-diff.Rmd
made compound dapc figure

```{r amova-no-cc}
(res   <- poppr.amova(dat11cc, ~Region/Host/Year, dist = dat11cc.bruvo))
set.seed(999)
(res.p <- randtest(res, nrepet = 999))
```

@@

```{r dapc-region, fig.width = 7, fig.height = 7}
region.dapc <- dapc(dat11, strata(dat11)$Region, n.pca = 25, n.da = 14)
scatter(region.dapc, scree.pca = TRUE,
        bg = "grey80",
        col = funky(nlevels(strata(dat11)$Region)), 
        posi.pca = "topright"
        )
ggcompoplot(region.dapc, setPop(dat11, ~Region), pal = funky, cols = 2)
```

>>>>

```{r amova-no-cc, cache = TRUE}
(res   <- poppr.amova(dat11cc, ~Region/Host/Year, dist = dat11cc.bruvo))
set.seed(999)
(res.p <- randtest(res, nrepet = 999))
```

@@

```{r dapc-region-calc, fig.width = 8, fig.height = 7, cache = TRUE}
set.seed(2017-07-05)
region.xval <- xvalDapc(tab(dat11cc), strata(dat11cc)$Region, n.pca = 15:25, 
                        n.rep = 100, parallel = "multicore", ncpus = 4)
region.dapc <- region.xval$DAPC
```

```{r dapc-region, fig.width = 9, fig.height = 8, cache = TRUE}
scatter(region.dapc, 
        scree.pca = TRUE,
        bg = "grey95",
        col = "black",
        # col = rev(funky(nlevels(strata(dat11cc)$Region))), 
        pch = seq(nlevels(strata(dat11cc)$Region)),
        legend = "true",
        posi.leg = "topleft",
        posi.pca = "topright"
        )
ggcompoplot(region.dapc, setPop(dat11cc, ~Region), pal = rev(funky(nlevels(strata(dat11cc)$Region))), cols = 2)
```

```{r fig.width = 10, fig.height = 6}
LDS <- bind_cols(Population = region.dapc$grp, as.data.frame(region.dapc$ind.coord)) %>%
  as_tibble()
LDS_pop <- LDS %>% 
  group_by(Population) %>% 
  summarize_all(mean) %>%
  rename_all(function(x) gsub("LD", "mean", x))
LDS <- full_join(LDS, LDS_pop)
LDS_PLOT <- ggplot(LDS, aes(x = LD1, y = LD2, color = Population)) + 
  geom_point(aes(fill = Population), alpha = 0.5, pch = 21, color = "black") +
  geom_segment(aes(x = mean1, y = mean2, xend = LD1, yend = LD2), alpha = 0.5) +
  stat_ellipse(type = "norm", level = 0.66, alpha = 0.75) + 
  ggrepel::geom_label_repel(aes(x = mean1, y = mean2, label = Population), 
                            data = LDS_pop, show.legend = FALSE, color = "black") +
  theme_bw() +  
  theme(aspect.ratio = 1/1.618) +
  theme(legend.position = "bottom") +
  theme(axis.text = element_blank()) + 
  theme(axis.title = element_blank()) + 
  theme(axis.ticks = element_blank()) + 
  viridis::scale_color_viridis(discrete = TRUE) +
  viridis::scale_fill_viridis(discrete = TRUE) +
  scale_y_continuous(breaks = 0) + 
  scale_x_continuous(breaks = 0) + 
  theme(panel.background = element_rect(fill = "grey95")) +
  theme(panel.grid.major = element_line(color = "black"))
LDS_PLOT
LDS_PLOT + facet_wrap(~Population)
```

Another way to visualize this is to look at the probability of assignment by
source pouplation.

```{r dapc-region-prob, fig.width = 7, fig.height = 6}
region_summary <- region.dapc$posterior %>% 
  as.data.frame() %>% 
  as_tibble() %>% 
  add_column(Individual = indNames(dat11cc)) %>% 
  add_column(`Original Population` = as.character(strata(dat11cc)$Region)) %>% 
  group_by(`Original Population`) %>%
  mutate(n = n()) %>%
  gather(Population, Posterior, -Individual, -`Original Population`, -n) %>% 
  group_by(Population, `Original Population`, n) %>% 
  summarize(`Mean Assignment` = mean(Posterior)) %>% 
  arrange(`Mean Assignment`) %>% 
  mutate(Source = case_when(
    Population == `Original Population` ~ "Original Population",
    TRUE ~ "Other"
    )) %>%
  ungroup() %>%
  arrange(desc(n)) %>%
  mutate(`Original Population` = forcats::fct_inorder(`Original Population`)) %>%
  mutate(Population = factor(Population, levels(`Original Population`))) %>%
  mutate(`Original Population` = paste0(`Original Population`, " (", n, ")")) %>%
  mutate(`Original Population` = forcats::fct_inorder(`Original Population`))
REGION_PLOT <- region_summary %>%
  filter(n >= 10) %>%
  ggplot(aes(x = Population, y = `Mean Assignment`)) +
  geom_segment(aes(xend = Population, yend = 0)) +
  geom_point(aes(fill = Source), pch = 21) + 
  facet_wrap(~`Original Population`, ncol = 3, strip.position = "top") +
  scale_fill_grey() +
  # scale_x_discrete(position = "top") +
  scale_y_continuous(limits = c(0, 1), minor_breaks = c(seq(0, 1, by = 0.125)), breaks = c(0, 0.5, 1)) +
  theme_bw(base_size = 16, base_family = "Helvetica") +
  theme(legend.position = "top") +
  theme(aspect.ratio = 1/2) +
  theme(axis.text = element_text(color = "black")) +
  theme(axis.ticks.y = element_blank()) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  # theme(axis.text.x.top = element_text(angle = 90, vjust = 0.5, hjust = 0)) +
  theme(panel.grid.major.x = element_line(linetype = 0, color = "grey50")) +
  theme(panel.grid.major = element_line(colour = "grey20")) +
  theme(panel.grid.minor = element_line(linetype = 3, colour = "grey50")) +
  # theme(panel.spacing.y = unit(0, "line")) +
  theme(strip.background = element_rect(color = NA, fill = "grey90")) +
  theme(strip.text = element_text(face = "bold", hjust = 0.05)) +
  theme(panel.border = element_blank()) +
  xlab("Population Assignment") +
  ylab("Mean Probability")
REGION_PLOT

```

```{r dapc-region-dualplot, fig.width = 13, fig.height = 6}
cowplot::plot_grid(REGION_PLOT, LDS_PLOT, 
                   nrow = 1, 
                   rel_widths = c(1, 0.85),
                   label_size = 16,
                   labels = "auto")
ggsave(filename = "results/figures/publication/DAPC.pdf")
```
